---
title: "davis_VAR_testing2"
output: html_document
date: "2023-08-16"
---

```{r, echo= FALSE}
library(tidyverse)
library(vars)
library(ggplot2)
library(forecast)
library(tseries)
library(gridExtra)
library(dplyr)
library(tidyr)

library(grDevices)
```

# load data set

```{r, echo=TRUE}
setwd("C:/Users/User/Desktop/R_Data_Science/cementresearch/")
df <- read_csv("data/combinedata6.csv") #includes real GDP
# df <- read_csv("data/combinedata5.csv")

variables <- c("Cement","unemploy","housing", "limestone","gdp")

start_year <- 1973 #we are working with 1971 - 2021 data now

df <- df |>
  filter(year >= start_year)


```

Variables considered: Cement, unemploy, limestone, gdp , housing

# adf test

```{r}

# ----------- SYNTHESIZE RESULTS -----------------

adf_results <- data.frame(variable = character(0), 
                          first_pvalue = character(0),
                          first_stationarity = character(0),
                          last_pvalue = character(0),
                          last_stationarity = character(0),
                          num_differences = integer(0))

# Loop through each variable
for (var in variables) {
  data <- df[[var]]  # Replace with your actual data extraction
  
  # Perform ADF test for the initial series
  adf_result_first <- adf.test(data)
  p_value_first <- adf_result_first$p.value
  formatted_p_value_first <- ifelse(p_value_first <= 0.01, "<0.01", sprintf("%.3f", p_value_first))
  stationarity_first <- ifelse(p_value_first <= 0.05, "Stationary", "Non-Stationary")
  
  num_diff <- 0
  p_value_last <- p_value_first
  stationarity_last <- stationarity_first
  
  # Iterate until achieving stationarity
  while (p_value_last > 0.05) {
    num_diff <- num_diff + 1
    data <- diff(data)
    adf_result_last <- adf.test(data)
    p_value_last <- adf_result_last$p.value
    formatted_p_value_last <- ifelse(p_value_last <= 0.01, "<0.01", sprintf("%.3f", p_value_last))
    stationarity_last <- ifelse(p_value_last <= 0.05, "Stationary", "Non-Stationary")
  }
  
  # Append results to the dataframe
  adf_results <- adf_results %>%
    add_row(variable = var,
            first_pvalue = formatted_p_value_first,
            first_stationarity = stationarity_first,
            last_pvalue = formatted_p_value_last,
            last_stationarity = stationarity_last,
            num_differences = num_diff)
}

# Display the rounded p-values results
print(adf_results)
```


# Adjusting for differencing 
```{r}
diff_1_vars_lim <- df[,c("unemploy","housing","Cement","gdp")]
diff_1_vars_lim <- ts(diff_1_vars_lim, start = start_year, frequency = 1)
diff_1_vars_lim <- as.data.frame(lapply(diff_1_vars_lim,function(x) diff(x)))

diff_2_vars_lim <- df[,c("limestone")]
diff_2_vars_lim <- ts(diff_2_vars_lim, start = start_year, frequency = 1)
diff_2_vars_lim <- as.data.frame(lapply(diff_2_vars_lim,function(x) diff(x, differences = 2)))

dummy_var <- data.frame(year = 1973:2021) |>
  mutate(value = case_when(year %in% c(1974,1975,1982,1992,1993, 2008,2009,2010) ~ 1,TRUE ~ 0))

differenced_df_lim <- cbind(
  diff_1_vars_lim[-1,],
  diff_2_vars_lim
)

differenced_df_lim <- ts(differenced_df_lim, start = start_year + 2,frequency = 1)

VARselect(differenced_df_lim,lag.max = 6,type = "trend",season = NULL, exogen = NULL) #7 is lag.max because 7^2 is 49 which is less than our 52 observations
 #BIC / SC Test

var_model_lim <- VAR(differenced_df_lim, p = 1, type = "trend") # need to justify why we can use the "both" parameter

forecast_lim <- predict(var_model_lim, n.ahead = 9,ci = 0.95)
par(col = "black", col.axis = "black",mfrow = c(4,1), mar = c(2,2,2,2))
plot(forecast_lim)
fanchart(forecast_lim)


tmin <- as.Date("1973-01-01")
tmax <- as.Date("2050-01-01")
tlab <- seq(tmin, tmax, by="year")
time <- substr(tlab, 0, 4)
filtered_years <- replace(time, !(time %in% c("1975", "1980", "1985", "1990", "1995", "2000", "2005", "2010", "2015", "2020", "2025", "2030", "2035", "2040", "2045", "2050")), NA)


fanchart2(forecast_lim
          ,xaxt="n" #removes original xaxis
          ,ylab = c("Thousand Metric Tons","Percent (%)","Thousands of Units","Thousand Metric Tons","Billions of USD") #### question 1: see hacked function ###
          ,main = c("Cement - 1st Difference","Unemployment - 1st Difference","Housing - 1st Difference", "Limestone - 2nd Difference","Real GDP - 1st Difference")
          ,colors= grDevices::gray.colors(5)#original is heat.colors(12)
          ,lab.at = seq(1,length(tlab)+0,1)
          ,lab.text = filtered_years
)


```

```{r}
# --------------------
forecast_original_lim_Cement <- cumsum(forecast_lim$fcst$Cement[,1]) + df$Cement[length(df$Cement)]
forecast_original_lim_housing <- cumsum(forecast_lim$fcst$housing[,1]) + df$housing[length(df$housing)]
forecast_original_lim_unemploy <- cumsum(forecast_lim$fcst$unemploy[,1]) + df$unemploy[length(df$unemploy)]
forecast_original_lim_limestone <- cumsum(cumsum(forecast_lim$fcst$limestone[,1]) + diff(df$limestone)[length(diff(df$limestone))]) + df$limestone[length(df$limestone)]
forecast_original_lim_gdp <- cumsum(forecast_lim$fcst$gdp[,1]) + df$gdp[length(df$gdp)]


forecast_original_lim_Cement
forecast_original_lim_housing
forecast_original_lim_unemploy
forecast_original_lim_limestone
forecast_original_lim_gdp


fcst_original_all <- bind_cols(data.frame(seq(2022,2030)),
                               bind_cols(forecast_original_lim_Cement,
                                         forecast_original_lim_unemploy,
                                         forecast_original_lim_housing,
                                         forecast_original_lim_limestone,
                                         forecast_original_lim_gdp)
                               )
colnames(fcst_original_all) <- c("year",variables) 

hist_df <- df[,c("year",variables)]


hist_fcst_all <- bind_rows(hist_df,fcst_original_all)

plot_val <- hist_fcst_all |>
  pivot_longer(cols = -c(year,Cement), names_to = "variable", values_to = "value")

plot_facet <- ggplot(plot_val, aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 2) +
  labs(x = "Years", y = "Value")

plot_cement <- ggplot(plot_val, aes(x = year, y = Cement)) + 
  geom_line() +
  xlab("Years") + ylab("Production (Mmt)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  scale_x_continuous(breaks = seq(1970, 2030, by = 5)) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") +
  ggtitle("Cement Production Projection in the U.S.") +
  theme_bw()

plot_cement 

plot_facet

```

#STEPWISE STUFF
```{r}
#Value of stepwise
intercept_coeff <- 2.465e+04
unemploy_coeff <- -1.043e+03
housing_coeff <- 1.009e+01
limestone_coeff <- 3.922e-01
gdp_coeff <- 7.934e-01


stepwise_vals <- bind_cols(data.frame(seq(start_year,2030), 
                                       data.frame(hist_fcst_all$unemploy * unemploy_coeff + hist_fcst_all$housing * housing_coeff + hist_fcst_all$limestone * limestone_coeff + hist_fcst_all$gdp * gdp_coeff + intercept_coeff)))
colnames(stepwise_vals) <- c("year","Cement_stepwise") 


stepwise_vals |>
  ggplot(aes(x = year, y= Cement_stepwise)) + 
  geom_line() +
  xlab("Years") + ylab("Production (Mmt)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  scale_x_continuous(breaks = seq(1970, 2030, by = 5)) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") +
  ggtitle("Cement Stepwise Production Projection in the U.S.") +
  theme_bw()

VAR_vs_stepwise <- bind_cols(hist_fcst_all[,c("year","Cement")],
                            stepwise_vals[,"Cement_stepwise"])
colnames(VAR_vs_stepwise) <- c("year","Cement_VAR","Cement_stepwise") 


VAR_vs_stepwise_new <- VAR_vs_stepwise
VAR_vs_stepwise_new$Cement_VAR_only_fcst <- ifelse(VAR_vs_stepwise_new$year < 2021, NA, VAR_vs_stepwise_new$Cement_VAR)
VAR_vs_stepwise_new$Cement_hist_only <- ifelse(VAR_vs_stepwise_new$year <= 2020, VAR_vs_stepwise_new$Cement_VAR, NA)

VAR_vs_stepwise_new |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement_hist_only, color = "1968-2021 Recorded Values")) +
  geom_line(aes(y = Cement_VAR_only_fcst, color = "VAR 2022-2050 Forecasts")) +
  geom_line(aes(y = Cement_stepwise, color = "Stepwise 1968-2050 Compared")) +
  xlab("Years") + ylab("Production (Mmt)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  scale_x_continuous(breaks = seq(1970, 2050, by = 5)) +
  scale_color_manual(values = c("VAR 2022-2050 Forecasts" = "#3e649e", "1968-2021 Recorded Values" = "black", "Stepwise 1968-2050 Compared" = "#00ba38")) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") +
  ggtitle("Cement VAR vs Stepwise Production Projection in the U.S.") +
  theme_bw()

```


#Each Individual Projection in its original units 

```{r}

# Splitting Data into forecasted value, and lower and upper bounds 
fcst_val <- as.data.frame(forecast_lim$fcst)[, grepl("fcst", colnames(as.data.frame(forecast_lim$fcst)))]
fcst_lower <- as.data.frame(forecast_lim$fcst)[, grepl("lower", colnames(as.data.frame(forecast_lim$fcst)))]
fcst_upper <- as.data.frame(forecast_lim$fcst)[, grepl("upper", colnames(as.data.frame(forecast_lim$fcst)))]

#Getting Real Values
consts_df <- as.data.frame(df[length(df$Cement),variables])

fcst_original_val <- fcst_val |>
  mutate_all(cumsum) + consts_df[rep(seq_len(nrow(consts_df)), each = nrow(fcst_val)), ]
fcst_original_val <- bind_cols(data.frame(seq(2022,2050)),fcst_original_val)
colnames(fcst_original_val) <- c("year",variables) 


fcst_original_lower <- fcst_lower |>
  mutate_all(cumsum) + consts_df[rep(seq_len(nrow(consts_df)), each = nrow(fcst_lower)), ]
fcst_original_lower <- bind_cols(data.frame(seq(2022,2050)),fcst_original_lower)
colnames(fcst_original_lower) <- c("year",variables) 


fcst_original_upper <- fcst_upper |>
  mutate_all(cumsum) + consts_df[rep(seq_len(nrow(consts_df)), each = nrow(fcst_upper)), ]
fcst_original_upper <- bind_cols(data.frame(seq(2022,2050)),fcst_original_upper)
colnames(fcst_original_upper) <- c("year",variables) 

#Putting together historical (hist) and forecasted data in original units (fcst)

hist_df <- df[,c("year",variables)]


hist_fcst_val <- bind_rows(hist_df,fcst_original_val)
hist_fcst_lower <- bind_rows(hist_df,fcst_original_lower)
hist_fcst_upper <- bind_rows(hist_df,fcst_original_upper)

plot_val <- hist_fcst_val |>
  pivot_longer(cols = -c(year,Cement), names_to = "variable", values_to = "value")

plot_facet <- ggplot(plot_val, aes(x = year, y = value)) +
  geom_line() +
  facet_wrap(~ variable, scales = "free_y", ncol = 2) +
  labs(x = "Years", y = "Value")

plot_cement <- ggplot(plot_val, aes(x = year, y = Cement)) + 
  geom_line() +
  xlab("Years") + ylab("Production (Mmt)") +
  scale_y_continuous(labels = scales::comma_format(scale = 1e-3)) +
  scale_x_continuous(breaks = seq(1970, 2050, by = 5)) +
  geom_vline(xintercept = 2021, color = "red", linetype = "dashed") +
  ggtitle("Cement Production Projection in the U.S.") +
  theme_bw()

plot_cement 

plot_facet

```

# Diagnostic Tests

```{r}
# Testing Autocorrelations/Serial Correlation 

mdl.serial <- serial.test(var_model_lim,lags.pt = 12,type="PT.asymptotic") #Use 12 because its like the default?

plot(mdl.serial, names = "Cement")

mdl.serial$resid

#Heteroskedasity - PASS, we fail to reject the null hypothesis so we conclude that there is not enough evidence to deduce heteroskedasity
mdl.arch <- arch.test(var_model_lim,lags.multi = 12,multivariate.only = TRUE)
plot(mdl.arch)

normality.test(var_model_lim,multivariate.only = TRUE)

mdl.norm
# Normality - Not sure if we can use normality here
par( mfrow= c(2,2) )
resdulpdf<-density(residuals[,"unemploy"])
plot(resdulpdf,main='Unemploy Residual shape Plot',xlab='Unemploy Residuals')
resdulpdf<-density(residuals[,"gdp"])
plot(resdulpdf,main='GDP Residual shape Plot',xlab='GDP Residuals')
resdulpdf<-density(residuals[,"housing"])
plot(resdulpdf,main='housing Residual shape Plot',xlab='housing Residuals')
resdulpdf<-density(residuals[,"limestone"])
plot(resdulpdf,main='limestone Residual shape Plot',xlab='limestone Residuals')




#Granger Causality - PASSS
granger_causality(var_model_lim)

causality_test_result$Granger 

#Impulse Response 
irf_results <- irf(var_model_lim, n.ahead = 9, response = c("unemploy", "gdp","housing","limestone"))

plot(irf_results)


#Stability
stability_result <- stability(var_model_lim,type = "OLS-CUSUM")

plot(stability_result)
roots(var_model_lim,modulus = TRUE)
sctest(var_model_lim)

#Correlation Matrix
summary(var_model_lim)
stargazer(var_model_lim[["varresult"]], type = "text")

```

#Plotting 

Per Witcraft request we need to get:

1) a plot with double axis with the x axis being years, the right y axisaxis being the GDP scale, and the left axis being the cement scale. We should have 3 lines on the plot, historical recorded cement, cement during historical times using VAR projections, and regular historical GDP 

2) Need to get a QQPLOT which is basically Cement projected by VAR vs recorded cement during the historical times 

THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR
THIS ONE PRODUCES FAULTY RESULTS BECAUSE IT ADDS RESIDUAL_DIF WITH REGULAR

```{r}

#getting the values of VAR predictived

df_hold <- df[,c("year",variables)] |>
  filter(year >= start_year+3) 

VAR_resid <- residuals(var_model_lim)

VAR_resid <- data.frame(year = seq(1974, 2021), VAR_resid)
VAR_resid <- VAR_resid[names(df_hold)]


VAR_hist_predict <- bind_cols(df_hold[1], df_hold[-1] + VAR_resid[-1])
colnames(VAR_hist_predict)[-1] <- paste(variables, "_VAR_predict",sep = "")

VAR_hist_predict_andrecorded <- bind_cols(VAR_hist_predict, df_hold[-1])
# Getting the first plot

scale = median(VAR_hist_predict_andrecorded$Cement_VAR_predict/VAR_hist_predict_andrecorded$gdp)
VAR_hist_predict_andrecorded |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_predict, color = "Cement_VAR_predict")) + 
  geom_line(aes(y = gdp*scale, color = "GDP")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_predict"="red","GDP" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="GDP")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 


scale = median(VAR_hist_predict_andrecorded$Cement_VAR_predict/VAR_hist_predict_andrecorded$housing)
VAR_hist_predict_andrecorded |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_predict, color = "Cement_VAR_predict")) + 
  geom_line(aes(y = housing*scale, color = "housing")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_predict"="red","housing" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="housing")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 

# Second plot

VAR_hist_predict_andrecorded |>
  ggplot(aes(x = Cement, y = Cement_VAR_predict)) + 
  geom_point() +
  geom_abline(slope=1, intercept=0) + 
  xlab("Data Source Recorded Cement") +ylab("VAR Predicted Cement")

qqnorm(VAR_resid$Cement)
qqline(VAR_resid$Cement)



shapiro.test(VAR_resid$Cement)

shapiro_result <- lapply(VAR_resid,function(x) shapiro.test(x))


shapiro_result
```


```{r}

df_hold <- df[,c("year",variables)] |>
  filter(year >= start_year+3) 

#-----------------------------using the fitted() method-----------------------------

VAR_fit <- fitted(var_model_lim)

all_Cement <- bind_cols(year = df_hold$year,Cement_real = df_hold$Cement, Cement_real_dif = diff(df$Cement)[-c(1,2)] , Cement_VAR_pred_dif = VAR_fit[,"Cement"])

all_Cement$Cement_VAR_pred <- cumsum(all_Cement$Cement_VAR_pred_dif) + df$Cement[df$year == 1973]

FITTED_METHOD_YES_DIF <- all_Cement |>
  ggplot(aes(x = year)) +
  geom_line(aes(y = Cement_real_dif, color = "Cement_real_dif")) + 
  geom_line(aes(y = Cement_VAR_pred_dif, color ="Cement_VAR_pred_dif")) +
  scale_color_manual(values = c("Cement_real_dif" = "blue", "Cement_VAR_pred_dif" = "red")) +
  ggtitle("USING FITTED() METHOD - differenced")

FITTED_METHOD_NO_DIF <- all_Cement |>
  ggplot(aes(x = year)) +
  geom_line(aes(y = Cement_real, color = "Cement_real")) + 
  geom_line(aes(y = Cement_VAR_pred, color ="Cement_VAR_pred")) +
  scale_color_manual(values = c("Cement_real" = "blue", "Cement_VAR_pred" = "red"))+
  ggtitle("USING FITTED() METHOD - no differenced")


#-----------------------------using the residual() method-----------------------------
VAR_resid <- residuals(var_model_lim)
VAR_resid <- data.frame(year = seq(1974, 2021), VAR_resid)
VAR_resid <- VAR_resid[names(df_hold)]
colnames(VAR_resid)[-1] <- paste(variables, "_resid_dif",sep = "")

VAR_resid_all <- bind_cols(year = df_hold$year,Cement_real = df_hold$Cement,Cement_real_dif = diff(df$Cement)[-c(1,2)], VAR_resid[-1], Cement_VAR_resid_pred_dif = diff(df$Cement)[-c(1,2)] - VAR_resid$Cement_resid_dif, ) #the diff(df$Cement)[-c(1,2)] is from 1974 - 21. #VAR_resid_pred_dif is NOT the residuals, its the predicted differenced data based on the RESIDUALS method

VAR_resid_all$Cement_VAR_resid_pred <- cumsum(VAR_resid_all$Cement_VAR_resid_pred_dif) + df$Cement[df$year == 1973]

RESIDUAL_METHOD_YES_DIF <- VAR_resid_all |>
  ggplot(aes(x = year)) +
  geom_line(aes(y = Cement_real_dif, color = "Cement_real_dif")) + 
  geom_line(aes(y = Cement_VAR_resid_pred_dif, color ="Cement_VAR_resid_pred_dif")) +
  scale_color_manual(values = c("Cement_real_dif" = "blue", "Cement_VAR_resid_pred_dif" = "red")) +
  ggtitle("USING RESIDUAL() METHOD -differenced")


RESIDUAL_METHOD_NO_DIF <- VAR_resid_all |>
  ggplot(aes(x = year)) +
  geom_line(aes(y = Cement_real, color = "Cement_real")) + 
  geom_line(aes(y = Cement_VAR_resid_pred, color ="Cement_VAR_resid_pred")) +
  scale_color_manual(values = c("Cement_real" = "blue", "Cement_VAR_resid_pred" = "red")) +
  ggtitle("USING RESIDUAL() METHOD - no differenced")

grid.arrange(FITTED_METHOD_YES_DIF, FITTED_METHOD_NO_DIF, RESIDUAL_METHOD_YES_DIF, RESIDUAL_METHOD_NO_DIF, ncol = 2)

# -------------- WITCRAFT PLOTS --------------------

scale = mean(VAR_resid_all$Cement_real/df_hold$gdp)
VAR_resid_all |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement_real, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_resid_pred, color = "Cement_VAR_resid_pred")) + 
  geom_line(aes(y = df_hold$gdp*scale, color = "gdp")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_resid_pred"="red","gdp" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="gdp")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 


scale = mean(VAR_resid_all$Cement_real/df_hold$housing)
VAR_resid_all |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement_real, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_resid_pred, color = "Cement_VAR_resid_pred")) + 
  geom_line(aes(y = df_hold$housing*scale, color = "housing")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_resid_pred"="red","housing" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="housing")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 


scale = mean(VAR_resid_all$Cement_real/df_hold$unemploy)
VAR_resid_all |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement_real, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_resid_pred, color = "Cement_VAR_resid_pred")) + 
  geom_line(aes(y = df_hold$unemploy*scale, color = "unemploy")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_resid_pred"="red","unemploy" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="unemploy")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 

scale = mean(VAR_resid_all$Cement_real/df_hold$limestone)
VAR_resid_all |>
  ggplot(aes(x = year)) + 
  geom_line(aes(y = Cement_real, color = "Cement_recorded")) +
  geom_line(aes(y = Cement_VAR_resid_pred, color = "Cement_VAR_resid_pred")) + 
  geom_line(aes(y = df_hold$limestone*scale, color = "limestone")) + 
  scale_color_manual(values = c("Cement_recorded" = "blue","Cement_VAR_resid_pred"="red","limestone" = "green")) +
  scale_y_continuous(sec.axis = sec_axis(~./scale, name="limestone")) +
  theme(axis.text.x = element_text(angle = 90)) + 
  ggtitle("VAR vs Recorded 1974 - 2021") +
  xlab("Years")+
  theme_bw() 


VAR_resid_all |>
  ggplot(aes(x = Cement_real, y= Cement_VAR_resid_pred)) +
  geom_point() + 
  geom_abline(slope=1, intercept=0) + 
  xlab("Data Source Recorded Cement") +ylab("VAR Predicted Cement")


lm_cement <- lm(Cement_real ~ Cement_VAR_resid_pred, data = VAR_resid_all)

summary(lm_cement)
plot(lm_cement)

```

#BEAUTIFUL HACKED VERSION OF vars::fanchart 

https://stackoverflow.com/questions/51601634/fanchart-customisation-vars-package
```{r}


fanchart2 <- function(x,
                       colors = NULL,
                       cis = NULL,
                       names = NULL,
                       main = NULL,
                       ylab = NULL,
                       xlab = NULL,
                       col.y = NULL,
                       add.preds = NULL,
                       nc,
                       plot.type = c("multiple","single"),
                       mar = par("mar"),
                       oma = par("oma"),
                       lab.at=NULL,
                       lab.text=NULL,...) {
  if (!(class(x) == "varprd")){
    stop("\nPlease provide an object of class 'varprd',\ngenerated by predict-method for objects of class 'varest'.\n")
  }
  if (is.null(colors)){
    colors <- gray(sqrt(seq(from = 0.05, to = 1, length = 9)))
  }


  if (is.null(cis)) {
    cis <- seq(0.1, 0.9, by = 0.1)
  }
  else {
    if ((min(cis) <= 0) || (max(cis) >= 1)){
      stop("\nValues of confidence intervals must be in(0, 1).\n")
    }
    if (length(cis) > length(colors)){
      stop("\nSize of 'colors' vector must be at least as long as\nsize of 'cis' vector\n")
    }
  }
  n.regions <- length(cis)
  n.ahead <- nrow(x$fcst[[1]])
  K <- ncol(x$endog)
  e.sample <- nrow(x$endog)
  endog <- x$endog
  fcst <- NULL
  for (j in 1:n.regions) {
    fcst[[j]] <- predict(x$model, n.ahead = n.ahead, ci = cis[j],
                         dumvar = x$exo.fcst)$fcst
  }
  xx <- seq(e.sample, length.out = n.ahead + 1)
  xx <- c(xx, rev(xx))
  op <- par(no.readonly = TRUE)
  plot.type <- match.arg(plot.type)
  ynames <- colnames(endog)
  if (is.null(names)) {
    names <- ynames
  }
  else {
    names <- as.character(names)
    if (!(all(names %in% ynames))) {
      warning("\nInvalid variable name(s) supplied, using first variable.\n")
      names <- ynames[1]
    }
  }
  nv <- length(names)
  ifelse(is.null(main), main <- paste("Fanchart for variable",
                                      names), main <- rep(main, nv)[1:nv])
  ifelse(is.null(ylab), ylab <- "", ylab <- ylab)
  ifelse(is.null(xlab), xlab <- "", xlab <- xlab)
  ifelse(is.null(col.y), col.y <- "black", col.y <- col.y)
  if (plot.type == "single") {
    if (nv > 1)
      par(ask = TRUE)
    par(mar = mar, oma = oma)
  }
  else if (plot.type == "multiple") {
    if (missing(nc)) {
      nc <- ifelse(nv > 4, 2, 1)
    }
    nr <- ceiling(nv/nc)
    par(mfcol = c(nr, nc), mar = mar, oma = oma)
  }
  for (i in 1:nv) {
    ymax <- max(c(fcst[[n.regions]][names[i]][[1]][, 3]),
                endog[, names[i]])
    ymin <- min(c(fcst[[n.regions]][names[i]][[1]][, 2]),
                endog[, names[i]])
    yy1 <- c(endog[e.sample, names[i]], fcst[[1]][names[i]][[1]][,
                                                                 2], rev(c(endog[e.sample, names[i]], fcst[[1]][names[i]][[1]][,
                                                                                                                               3])))
    plot.ts(c(endog[, names[i]], rep(NA, n.ahead)),
            main = main[i],
            ylim = c(ymin, ymax),
            ylab = ylab[i],#### question 1 #### modivied ylab to depend on the loop counter
            xlab = xlab,
            col = col.y,
            ...)
    polygon(xx, yy1, col = colors[1], border = colors[1])
    if (n.regions > 1) {
      for (l in 2:n.regions) {
        yyu <- c(endog[e.sample, names[i]], fcst[[l]][names[i]][[1]][,
                                                                     3], rev(c(endog[e.sample, names[i]], fcst[[l -
                                                                                                                  1]][names[i]][[1]][, 3])))
        yyl <- c(endog[e.sample, names[i]], fcst[[l -
                                                    1]][names[i]][[1]][, 2], rev(c(endog[e.sample,
                                                                                         names[i]], fcst[[l]][names[i]][[1]][, 2])))
        polygon(xx, yyu, col = colors[l], border = colors[l])
        polygon(xx, yyl, col = colors[l], border = colors[l])
      }
    }

    #### question 4 ####
    # if a matrix of points at various times in the prediction is sent to the function
    # they will be plotted here
    # standard adjustments to color and pch are possible
    # assumes a matrix of values is given with columns in order of the variables (price=col 1)
    # and NA values are times without prediction and not plotted
    if(is.null(add.preds)==F){
      points(x=xx[2:(length(xx)/2)]
             ,y=add.preds[,i]
             ,col='gray48'
             ,pch=16)
    }

    #### question 3: adding axis to each plot inside the hacked function ####
    # standard modifications can be done to this function
    if(is.null(lab.at)==F){
      axis(1,
           at=lab.at,
           labels=lab.text,
           las=3,
           line=1,
           cex.axis=1) #orignal is 0.6
    }

  }
  on.exit(par(op))
}
```
